Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-24 src/2.GenericProgramming.tex
Changes to accodate semantic capture.

A programming model exposes constrained semantics to the programmer which the programmer uses to express intent. Absent an embedded programming model, the semantics of C-like languages can lead to sequential thinking and execution. Parallel programming models expose parallel semantics to enable developers to create performance portable code. This is commonly achieved by adding semantic value to the base programming language, often referred to as programming with \emph{pragma} annotations. Other methods expose such semantics through library calls or new languages. All three approaches are conceptually suitable to represent programming paradigms such as tasking, parallel patterns and others. Some representatives are the OpenMP programming model~\cite{CITEOPENMP}, threading libraries, or languages like Erlang in which parallelism is a first-class citizen. In general, it is up to the parallel programming model to expose language constructs or library interfaces (APIs) such that they combine convenience to the programmer with semantics sufficient to map the program execution to the parallel computer hardware efficiently and correctly. Questions of which semantic information is needed, what paradigm to present and by what syntactical means span an exciting design space worth exploring. From a theoritecal perspective

 Figure~\ref{figSemCapture} shows the three  areas that constitute a parallel programming model and its design. We call it the semantic capture.

\begin{figure}[h]
\begin{Verbatim}[frame=leftline]
Constrained Semantics: memory allocations, memory movement, parallel kernels
Parallel Paradigms: patterns, tasks
Expression: language 
Concern: prescriptive, descriptive
\end{Verbatim}
\caption{Semantic capture - a triplet of constrained semantics, parallel paradigms, and expression define by a parallel programming model to allow developers to run performance-portable code on a set of architectures}
\label{figSemCapture}
\end{figure}

>>>>

A programming model exposes constrained semantics to the programmer which the programmer uses to express intent. Absent an embedded programming model, the semantics of C-like languages can lead to sequential thinking and execution. Parallel programming models expose parallel semantics to enable developers to create performance portable code. This is commonly achieved by adding semantic value to the base programming language, often referred to as programming with \emph{pragma} annotations. Other methods expose such semantics through library calls or new languages. All three approaches are conceptually suitable to represent programming paradigms such as tasking, parallel patterns and others. Some representatives are the OpenMP programming model~\cite{CITEOPENMP}, threading libraries, or languages like Erlang in which parallelism is a first-class citizen. In general, it is up to the parallel programming model to expose language constructs or library interfaces (APIs) such that they combine convenience to the programmer with semantics sufficient to map the program execution to the parallel computer hardware efficiently and correctly. Questions of which semantic information is needed, what paradigm to present and by what syntactical means span an exciting design space worth exploring.

 Figure~\ref{figSemCapture} shows the four properties that constitute a parallel programming model and its design. We call it the semantic capture.

\begin{figure}[h]
\begin{Verbatim}[frame=leftline]
Constrained Semantics: memory allocations, memory movement, parallel kernels
Parallel Paradigms: patterns, tasks
Expression: language 
Concern: prescriptive, descriptive
\end{Verbatim}
\caption{Semantic capture - a set of constrained semantics, parallel paradigms, expression and concern define key properties a parallel programming model.}
\label{figSemCapture}
\end{figure}

### 2020-01-24 src/4.ProgammingModel.tex
Fused DZP and JC changes for Chapters 1 and 2.
Changes to accodate semantic capture.

The Kokkos programming model specifies a language, programming paradigm and an API. The API exposes the abstract machine model to the developer and defines a set of pattens and types that capture intents and properties of execution. This completes the semantic capture, discussed in Chapter~\ref{chap:background}, We detail all three components as follows.

@@

\begin{figure}[h]
\begin{Verbatim}[frame=leftline]
Semantic information: Patterns (intent), Spaces, Layouts, 
Policies and Traits
Paradigm: Parallel patterns and tasking
Language: C++
\end{Verbatim}
\caption{Semantic capture defined by the Kokkos programming model.}
\label{fig:SemCaptureKokkos}
\end{figure}

>>>>

The Kokkos programming model specifies a language, programming paradigm and an API. The API exposes the machine representation to the developer and defines a set of pattens and types that capture intents and properties of execution. This completes the semantic capture, discussed in Chapter~\ref{chap:background}, We detail all three components as follows.

@@

\begin{figure}[h]
\begin{Verbatim}[frame=leftline]
Constrained Semantics: Patterns (intent), Spaces, Layouts, 
Policies and Traits
Parallel Paradigms: Parallel patterns and tasking
Expression: Embedded, C++
Concern: descriptive
\end{Verbatim}
\caption{Semantic capture defined by the Kokkos programming model.}
\label{fig:SemCaptureKokkos}
\end{figure}
